Report on the Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2017-03-09 WSSSPE4_report/WSSSPE4_report.tex
small fixes to new papers section and a todo

\begin{itemize}
\item \textbf{Author: Michael Heroux \emph{Idea paper: Sustainable & Productive: Improving Incentives for Quality Software}}.
The author presents that improving quality of scientific software is a measure  to address the increased complexity in both modeling and software. To support quality, it is essential to provide Computational Science and Engineering teams with publication venues, funding, and  professional recognition. Broad improvements in sustainability and productivity can be only achieved when publishers, funding  agencies  and  employers  raise  their  expectations for software quality. Additionally, software community leaders are in the unique position to positively impact software quality by working on establishing incentives that will spur creative and novel  approaches to improve developer productivity  and software sustainability.

\item \textbf{Author: Gabrielle Allen \emph{Idea Paper: Establishing a Professional Society for Research Software}}.
This paper advocates to create a professional society or extend the scope of an existing organization in research software. The author presents the motivation, a variety of potential roles of such a society and diverse potential funding mechanisms to be able to establish a sustainable organization. She gives an overview on similar activities as well as existing organizations and aims at stimulating discussion among potential members and stakeholders.

\item \textbf{Authors: Olivier Philippe, Simon Hettrick and Neil Chue Hong \emph{Experience Paper: Preliminary analysis of a survey of UK Research Software Engineers}}.
This paper presents results from a survey conducted on the new role of research software engineers in academia. The  survey  provides   demographic  information  about  the  education,  field,  gender,  job  satisfaction  and career  plans  of  the  people  in this community. the outcome is that members of the  community  is  found  to be highly educated, derive mainly from the hard sciences, and to be predominantly male. Respondents report satisfaction in their jobs,  but  indicate  that  career  progression  is  both  difficult  and opaque. The authors urge a  continued  discussion  about  the  experience  of research software engineers and  recommend further  investigation  into  this
important  community.

\item \textbf{Author: Christopher Gwilliams \emph{Demo: Using Industrial Engagement to Create and Develop Research Ties within Academia}}.
This demo focuses on the benefits of including a  new  software engineering degree that is influenced by, and works directly with,
industry.  Cardiff  University  introduced  a  new  industry-focused  degree, BSc Applied Software Engineering (ASE) in September
2015. The engagement with industry allows students to learn through live
projects  with  real-world  applications and forge links for career opportunities. Also industry benefits from this course that provides an entry point for industry to engage with the university and to develop a network of contacts between industry and academia, which is also a benefit for the university and allows to  identify  potential  opportunities for new research projects and collaborators.

\item \textbf{Authors: Anshu Dubey and Katherine Riley \emph{Experience Paper: Software Engineering and Community Codes Track in ATPESC}}.
The Argonne Training Program in Extreme Scale
Computing (ATPESC) was started by Argonne National Laboratory with the objective of expanding the ranks of better prepared users of high-performance computing (HPC) machines.
One of the unique aspects of the program was inclusion of a
software engineering and community codes track. The inclusion
was motivated by the observation that the projects with a good
software process were better able to meet their scientific goals.
The paper presents the experience in running the software
track, the motivation, reception, and evolution of the track over the years.  
The authors recommend the inclusion of similar tracks in other HPC-oriented training programs.

\item \textbf{Author: Neil Chue Hong \emph{Position Paper: Why do we need to compare research software, and how should we do it?}}.
This position paper sets out the reasons for why different  stakeholders, from users to developers 
to funders, might wish to undertake the difficult task to compare research software. The author describes a proposed framework for doing so (based around measures of accessibility, usability, maintainability and portability) which takes into account the possibility of variation between different communities about how they prioritise different aspects of research software.

\item \textbf{Authors: Anshu Dubey and Lois Curfman McInnes \emph{Idea Paper: The Lifecycle of Software for Scientific Simulation}}.
The authors discuss the specifics of the lifecycle of scientific software. While the software lifecycle is a well-researched topic, one class of projects, research software, has received relatively
little attention from lifecycle researchers. This
paper formalizes a lifecycle model
for end-to-end scientific application software, featuring loose
coupling betweensubmodels for development of infrastructure
and scientific capability. The authors also invite input from stakeholders
to converge on a model that captures the complexity of this
development processes and provides needed lifecycle guidance to
the scientific software community.

\item \textbf{Authors: Francisco Queiroz and Rejane Spitz \emph{Position Paper: Collaborative Gamification Design for Scientific Software}}. 
The paper focuses on usability design and presents that gamification, a design trend for improving scientific software usability, for scientific software should be facilitated by an open, collaborative design process supported by 
conversational media. This approach is compatible with 
qualities often attributed to computational science community 
regarding openness and collaboration between members of varied 
professional backgrounds. They exemplify the use of conversational media for collaborative design and expect the synergy between collaborators to result in better usability,  greater user acceptance, and adequacy to requirements, 
obtaining optimal design solutions in a sustainable way.

\item \textbf{Authors: Bruce Childers, Jack Davidson, Wayne Graves, Bernard Rous and David Wilkinson \emph{Position Paper: Active Curation of Artifacts is Changing the Way Digital Libraries will Operate}}.
The paper presents the implications of “active curation systems” that provide executable access to
artifacts and experiments behind published results and enables
their reuse. These systems allow changing and repeating experiments to understand how an innovation behaves in conditions
beyond the ones described in a paper and address reproducibility challenges. Active curation systems
also  enable accountability and accelerate research progress  by
giving access to complete experimental details. As these systems
take  hold,  it  is  important  to  understand  their  capabilities  and
how digital libraries should be integrated with them. The authors describe a study underway to explore how best
to do this integration.

\item \textbf{Authors: Debashis Ganguly, William C. Garrison III, David Wilkinson, Bruce R. Childers, Adam J. Lee, and Daniel Mosse \emph{Demo: Composing, Reproducing, and Sharing Simulations}}.
The authors demonstrate their approach of sharing,
reproducing, and composing simulations toward accelerating
research productivity while also improving accountability and
credibility. Specifically, they have developed a case study in
which they compose and share access control simulations in the
form of shareable data store units for cloud systems. They share, compose
and repeat simulations  through a collaborative repository
(OCCAM) and suggest a general simulation framework (SST) that can
accelerate the efforts for the whole community.

\item \textbf{Authors: Sameer Shende and Allen Malony. \emph{Demo: Using TAU for Performance Evaluation of Scientific Software}}.
This demo presents the  TAU Performance System for performance evaluation of Scientific 
Software written in C++, C, and Fortran. The system addresses performance 
technology problems  at three levels: instrumentation, 
measurement, and analysis. The TAU framework supports the 
configuration and integration of these layers. The authors advocate that effective 
exploration of performance will necessarily
require prudent selection from the range of alternative methods TAU provides 
to  assemble meaningful performance experiments.

\item \textbf{Authors: Janos Sallai, Christopher Iacovella, Christoph Klein and Tengyu Ma. \emph{Idea Paper: Development of a Software Framework for Formalizing Forcefield Atom-Typing for Molecular Simulation}}.

The authors present an essential and error-prone task in the area of Molecular Dynamics simulations: the efficient selection of correct parameter values for specific molecules, especially forcefields. They  present  a  framework  that  aims  to  solve  this
data  management  issue,  proposing  a  common  format  for
forcefields that is self-documenting with machine readable,
declarative usage rules. They investigate processes and tools
that are commonly used today in software development (e.g,
unit testing, verification and validation, continuous integra-
tion,  and  version  control)  and reason that they are, with proper infrastructure
support, applicable to forcefield development, as well. The
paper describes how such an infrastructure can tackle man-
aging and evolving forcefields by the MD community, and
proposes a way to encourage and incentivize involvement
by the stakeholders.

>>>>

\todo{group by type - idea, position, experience, demo}
\begin{itemize}
\item \textbf{Michael Heroux \emph{Idea paper: Sustainable \& Productive: Improving Incentives for Quality Software}}.
The author presents that improving quality of scientific software is a measure  to address the increased complexity in both modeling and software. To support quality, it is essential to provide Computational Science and Engineering teams with publication venues, funding, and  professional recognition. Broad improvements in sustainability and productivity can be only achieved when publishers, funding  agencies  and  employers  raise  their  expectations for software quality. Additionally, software community leaders are in the unique position to positively impact software quality by working on establishing incentives that will spur creative and novel  approaches to improve developer productivity  and software sustainability.

\item \textbf{Gabrielle Allen \emph{Idea Paper: Establishing a Professional Society for Research Software}}.
This paper advocates to create a professional society or extend the scope of an existing organization in research software. The author presents the motivation, a variety of potential roles of such a society and diverse potential funding mechanisms to be able to establish a sustainable organization. She gives an overview on similar activities as well as existing organizations and aims at stimulating discussion among potential members and stakeholders.

\item \textbf{Olivier Philippe, Simon Hettrick and Neil Chue Hong \emph{Experience Paper: Preliminary analysis of a survey of UK Research Software Engineers}}.
This paper presents results from a survey conducted on the new role of research software engineers in academia. The  survey  provides   demographic  information  about  the  education,  field,  gender,  job  satisfaction  and career  plans  of  the  people  in this community. the outcome is that members of the  community  is  found  to be highly educated, derive mainly from the hard sciences, and to be predominantly male. Respondents report satisfaction in their jobs,  but  indicate  that  career  progression  is  both  difficult  and opaque. The authors urge a  continued  discussion  about  the  experience  of research software engineers and  recommend further  investigation  into  this
important  community.

\item \textbf{Christopher Gwilliams \emph{Demo: Using Industrial Engagement to Create and Develop Research Ties within Academia}}.
This demo focuses on the benefits of including a  new  software engineering degree that is influenced by, and works directly with,
industry.  Cardiff  University  introduced  a  new  industry-focused  degree, Bachelor of Science in Applied Software Engineering (ASE) in September
2015. The engagement with industry allows students to learn through live
projects  with  real-world  applications and forge links for career opportunities. Also industry benefits from this course that provides an entry point for industry to engage with the university and to develop a network of contacts between industry and academia, which is also a benefit for the university and allows to  identify  potential  opportunities for new research projects and collaborators.

\item \textbf{Anshu Dubey and Katherine Riley \emph{Experience Paper: Software Engineering and Community Codes Track in ATPESC}}.
The Argonne Training Program in Extreme Scale
Computing (ATPESC) was started by Argonne National Laboratory with the objective of expanding the ranks of better prepared users of high-performance computing (HPC) machines.
One of the unique aspects of the program was inclusion of a
software engineering and community codes track. The inclusion
was motivated by the observation that the projects with a good
software process were better able to meet their scientific goals.
The paper presents the experience in running the software
track, the motivation, reception, and evolution of the track over the years.  
The authors recommend the inclusion of similar tracks in other HPC-oriented training programs.

\item \textbf{Neil Chue Hong \emph{Position Paper: Why do we need to compare research software, and how should we do it?}}.
This position paper sets out the reasons for why different  stakeholders, from users to developers 
to funders, might wish to undertake the difficult task to compare research software. The author describes a proposed framework for doing so (based around measures of accessibility, usability, maintainability and portability) which takes into account the possibility of variation between different communities about how they prioritize different aspects of research software.

\item \textbf{Anshu Dubey and Lois Curfman McInnes \emph{Idea Paper: The Lifecycle of Software for Scientific Simulation}}.
The authors discuss the specifics of the lifecycle of scientific software. While the software lifecycle is a well-researched topic, one class of projects, research software, has received relatively
little attention from lifecycle researchers. This
paper formalizes a lifecycle model
for end-to-end scientific application software, featuring loose
coupling between submodels for development of infrastructure
and scientific capability. The authors also invite input from stakeholders
to converge on a model that captures the complexity of this
development processes and provides needed lifecycle guidance to
the scientific software community.

\item \textbf{Francisco Queiroz and Rejane Spitz \emph{Position Paper: Collaborative Gamification Design for Scientific Software}}. 
The paper focuses on usability design and presents that gamification, a design trend for improving scientific software usability, for scientific software should be facilitated by an open, collaborative design process supported by 
conversational media. This approach is compatible with 
qualities often attributed to computational science community 
regarding openness and collaboration between members of varied 
professional backgrounds. They exemplify the use of conversational media for collaborative design and expect the synergy between collaborators to result in better usability,  greater user acceptance, and adequacy to requirements, 
obtaining optimal design solutions in a sustainable way.

\item \textbf{Bruce Childers, Jack Davidson, Wayne Graves, Bernard Rous and David Wilkinson \emph{Position Paper: Active Curation of Artifacts is Changing the Way Digital Libraries will Operate}}.
The paper presents the implications of ``active curation systems'' that provide executable access to
artifacts and experiments behind published results and enables
their reuse. These systems allow changing and repeating experiments to understand how an innovation behaves in conditions
beyond the ones described in a paper and address reproducibility challenges. Active curation systems
also  enable accountability and accelerate research progress  by
giving access to complete experimental details. As these systems
take  hold,  it  is  important  to  understand  their  capabilities  and
how digital libraries should be integrated with them. The authors describe a study underway to explore how best
to do this integration.

\item \textbf{Debashis Ganguly, William C. Garrison III, David Wilkinson, Bruce R. Childers, Adam J. Lee, and Daniel Mosse \emph{Demo: Composing, Reproducing, and Sharing Simulations}}.
The authors demonstrate their approach of sharing,
reproducing, and composing simulations toward accelerating
research productivity while also improving accountability and
credibility. Specifically, they have developed a case study in
which they compose and share access control simulations in the
form of shareable data store units for cloud systems. They share, compose
and repeat simulations  through a collaborative repository
(OCCAM) and suggest a general simulation framework (SST) that can
accelerate the efforts for the whole community.

\item \textbf{Sameer Shende and Allen Malony. \emph{Demo: Using TAU for Performance Evaluation of Scientific Software}}.
This demo presents the  TAU Performance System for performance evaluation of Scientific 
Software written in C++, C, and Fortran. The system addresses performance 
technology problems  at three levels: instrumentation, 
measurement, and analysis. The TAU framework supports the 
configuration and integration of these layers. The authors advocate that effective 
exploration of performance will necessarily
require prudent selection from the range of alternative methods TAU provides 
to  assemble meaningful performance experiments.

\item \textbf{Janos Sallai, Christopher Iacovella, Christoph Klein and Tengyu Ma. \emph{Idea Paper: Development of a Software Framework for Formalizing Forcefield Atom-Typing for Molecular Simulation}}.

The authors present an essential and error-prone task in the area of Molecular Dynamics simulations: the efficient selection of correct parameter values for specific molecules, especially forcefields. They  present  a  framework  that  aims  to  solve  this
data  management  issue,  proposing  a  common  format  for
forcefields that is self-documenting with machine readable,
declarative usage rules. They investigate processes and tools
that are commonly used today in software development (e.g,
unit testing, verification and validation, continuous integration,
and  version  control)  and reason that they are, with proper infrastructure
support, applicable to forcefield development, as well. The
paper describes how such an infrastructure can tackle man-
aging and evolving forcefields by the MD community, and
proposes a way to encourage and incentivize involvement
by the stakeholders.
